%%
%% RegDivergence-101
%% Reformatted to the official ACM acmart "sigconf" template.
%%
%% For submission/review you may instead use:
%%   \documentclass[sigconf, review, anonymous]{acmart}
%% For the camera-ready, use:
%%   \documentclass[manuscript]{acmart}
%%
\documentclass[manuscript]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{Bib\TeX}}}

%% Rights — replace these SAMPLE values with the commands from your ACM
%% rights-confirmation email after acceptance.
\setcopyright{acmlicensed}
\copyrightyear{2026}
\acmYear{2026}
\acmDOI{XXXXXXX.XXXXXXX}
\acmConference[AIFM '26]{2026 2nd International Conference on Artificial
  Intelligence and Foundation Model}{June 26--28, 2026}{Urumqi, China}
\acmISBN{979-8-4007-2458-9}

%%
%% Submission ID — uncomment and fill in when assigned
%%\acmSubmissionID{XXX-XXXX}

%% Fit-to-margin helpers (single-column): allow long \texttt tokens to break
\usepackage[htt]{hyphenat}
\emergencystretch=2em

\begin{document}

%%
%% Title. Use \title[short title]{full title} when the full title is long.
%%
\title[RegDivergence-101]{RegDivergence-101: An LLM Benchmark for
  Cross-Jurisdiction Regulatory Contradiction Detection in Life Sciences}

%%
%% Authors
%%
\author{Chuchu Wu}
\authornote{Corresponding author.}
\email{chuchuw@alumni.cmu.edu}
\orcid{0009-0006-5140-2247}
\affiliation{%
  \institution{Carnegie Mellon University}
  \city{Pittsburgh}
  \state{PA}
  \country{USA}
}

\author{Zhiyin Zhou}
\email{zzhou20@pratt.edu}
\orcid{0009-0004-5139-4668}
\affiliation{%
  \institution{Pratt Institute}
  \city{New York}
  \state{NY}
  \country{USA}
}

\author{Jingzhuo Hu}
\email{jingzhuoh00@gmail.com}
\orcid{0009-0008-0206-8531}
\affiliation{%
  \institution{University of Pennsylvania}
  \city{Philadelphia}
  \state{PA}
  \country{USA}
}

\author{Liang You}
\email{liy121@pitt.edu}
\orcid{0009-0002-8773-0614}
\affiliation{%
  \institution{University of Pittsburgh}
  \city{Pittsburgh}
  \state{PA}
  \country{USA}
}

\renewcommand{\shortauthors}{Wu et al.}

%%
%% Abstract
%%
\begin{abstract}
Pharmaceutical sponsors developing a drug for both the United States and the
European Union must reconcile guidance issued independently by the FDA and the
EMA. Where the two agencies require \textit{substantively the same thing}, a
sponsor can file once; where they \textit{diverge}, a single trial design risks
rejection in one region; where one agency is \textit{silent} on a point the
other regulates, the sponsor must infer obligations. Today this reconciliation
is performed manually by regulatory-affairs experts. We introduce
\textbf{cross-jurisdiction regulatory divergence detection}: given an FDA
requirement and an EMA requirement on the same topic, classify their
relationship as AGREE, DIVERGE, or SILENT. SILENT is inherently directional
(SILENT\_FDA vs.\ SILENT\_EMA); we record direction per pair and report
per-direction F1 alongside the collapsed label. We release
\textbf{RegDivergence-101}, a 101-pair expert-grounded \textbf{pilot}
evaluation benchmark (labels grounded in three peer-reviewed FDA/EMA
comparison studies and primary FDA/EMA/ICH guidance text; dual-annotation
inter-annotator $\kappa = 0.85$), and systematically characterise a four-method
baseline hierarchy: lexical heuristic (0.511 macro-F1, 95\% CI [0.411--0.605]), NLI
cross-encoder (0.233), obligation-level Graph-RAG (0.663 [0.570--0.747]), and
flat LLM judge / Claude Haiku (0.830 [0.747--0.908]). Three directional
observations emerge at pilot scale ($n = 101$): SILENT is semantically
detectable but invisible to entailment-only formulations; pair-level obligation
graphs improve over lexical methods but trail flat-LLM context (CIs partially
overlapping); and corpus-level graph construction is the indicated
architectural target for large-scale silent-detection. RegDivergence-101 is a
pilot release establishing the task formulation and baseline hierarchy; four
unrepresented regulatory domains and an expansion roadmap are described in
\S\ref{sec:futurework}.
\end{abstract}

%%
%% CCS Concepts — generated from https://dl.acm.org/ccs
%%
\begin{CCSXML}
<ccs2012>
  <concept>
    <concept_id>10010147.10010257</concept_id>
    <concept_desc>Computing methodologies~Natural language processing</concept_desc>
    <concept_significance>500</concept_significance>
  </concept>
  <concept>
    <concept_id>10002951.10003317.10003347</concept_id>
    <concept_desc>Information systems~Document filtering</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
  <concept>
    <concept_id>10003456.10003457.10003527</concept_id>
    <concept_desc>Applied computing~Health informatics</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Natural language processing}
\ccsdesc[300]{Information systems~Document filtering}
\ccsdesc[300]{Applied computing~Health informatics}

%%
%% Keywords
%%
\keywords{large language models, LLM benchmark, regulatory NLP, AI for
  compliance, FDA/EMA divergence, contradiction detection, Graph-RAG,
  retrieval-augmented generation, generative AI, life sciences,
  AGREE/DIVERGE/SILENT classification}

\maketitle

%% -------------------------------------------------------------------------
\section{Introduction}
%% -------------------------------------------------------------------------

Bringing a medicine to market in both the US and EU requires satisfying two
regulators that publish guidance separately and frequently \textit{differently}.
Gene-therapy sponsors face a 15-year long-term follow-up recommendation from FDA
while EMA defers the duration to case-by-case risk assessment---same topic,
materially different specificity. Analogous divergences span pediatric
extrapolation, endpoint definitions in ulcerative colitis, biosimilar
comparability, and estimand implementation. A sponsor who misses one divergence
during protocol design can lose a year or more re-running a study.

Reconciling the two corpora is currently a manual, expert task. Yet the NLP
building blocks for automating it all exist---document-level NLI
(ContractNLI~\cite{koreeda2021contractnli}, DocNLI~\cite{yin2021docnli}),
legal contradiction detection (LegalWiz~\cite{legalwiz2025},
CLAUSE~\cite{betterclause2026}), and Graph-RAG (Microsoft
GraphRAG~\cite{edge2024graphrag}, GraphCompliance~\cite{graphcompliance2025},
RAGulating Compliance~\cite{ragulating2025}). \textbf{No prior work combines
them to detect divergence between two parallel regulatory corpora from different
jurisdictions.} All existing contradiction work is \textit{intra-document}
(within one contract) or \textit{single-jurisdiction} (contract-vs-statute).

This paper makes three contributions:
\begin{enumerate}
\item \textbf{Task formulation.} Cross-jurisdiction regulatory divergence
  detection as a three-way classification (AGREE / DIVERGE / SILENT) over
  pre-aligned FDA/EMA requirement pairs, grounded in published expert
  comparisons. The task is explicitly scoped to the \textit{classification}
  stage given pre-aligned pairs; topic alignment is a non-trivial upstream
  problem addressed in \S\ref{sec:taskdef} and \S\ref{sec:dataset}. SILENT
  is treated as directional (SILENT\_FDA vs.\ SILENT\_EMA); we report
  per-direction F1 in Table~\ref{tab:silent_direction}.
\item \textbf{Pilot benchmark.} \textbf{RegDivergence-101}, a pilot 101-pair
  benchmark (59 from peer-reviewed FDA/EMA comparison studies, 42 derived from
  primary FDA/EMA/ICH guidance) spanning 13 topic areas, with a documented
  two-track pair-construction protocol, inter-annotator agreement
  ($\kappa = 0.85$), and bootstrap confidence intervals. Four regulatory
  domains are currently unrepresented; \S\ref{sec:futurework} describes the
  expansion roadmap.
\item \textbf{Baseline hierarchy with three directional observations.} Four
  methods systematically compared at pilot scale ($n = 101$), yielding
  directional empirical observations on the task's structure and the path to a
  corpus-level solution. Observations are reported with bootstrap 95\% CIs;
  confirmatory significance testing awaits a larger benchmark.
\end{enumerate}

%% -------------------------------------------------------------------------
\section{Related Work}
%% -------------------------------------------------------------------------

\textbf{Document-level NLI.} ContractNLI~\cite{koreeda2021contractnli} introduced
document-level entailment/contradiction/not-mentioned classification with
evidence extraction, and showed that contradiction detection lags entailment,
especially under ``negation by exception.'' DocNLI~\cite{yin2021docnli} extended
NLI to full documents.

\textbf{Legal contradiction detection (2025--2026).}
LegalWiz~\cite{legalwiz2025} uses a multi-agent generation framework for legal
contradiction detection but is confined to single-document regimes. Better Call
CLAUSE~\cite{betterclause2026} benchmarks 7,500+
perturbed contracts across ten anomaly categories and finds that LLMs miss
subtle errors and struggle to \textit{justify} them legally. Contradiction
Detection in RAG Systems~\cite{contradictionrag2025} uses LLMs as context
validators. Two recent RAG-faithfulness studies bear directly on our LLM-judge
and graph-RAG results: \citet{chen2026ragwrong} probe whether a RAG system can
recognise when retrieved context is incorrect, and \citet{qian2026evidenceforce}
show that the surface relevance of retrieved evidence does not guarantee it
\textit{warrants} a claim---the same failure mode documented in our Observation~2, where
modal-register similarity (RECOMMENDED vs.\ MANDATORY) misleads the pair-level
graph.

\textbf{Graph-RAG and regulatory compliance.}
Microsoft GraphRAG~\cite{edge2024graphrag} builds entity graphs with community
summaries for local and global queries. GraphCompliance~\cite{graphcompliance2025}
aligns \textit{policy graphs} (regulatory requirements) with \textit{context
graphs} (organisational facts) for GDPR; it explicitly lists ``handling
conflicting regulatory interpretations'' as open. RAGulating
Compliance~\cite{ragulating2025} builds an ontology-free regulatory knowledge
graph for traceable QA.

\textbf{Cross-jurisdiction regulatory benchmarks.}
The closest concurrent work is \textbf{Sino-US-DrugQA}~\cite{sinousdrugqa2026}
(11{,}871 MCQA items, US--CN), which shares our concept-alignment thesis but
cannot express the SILENT class. Comparative-regulatory-science surveys document
FDA/EMA divergence across dosing~\cite{mita2026doses}, oncology
labelling~\cite{gomezfernandez2026inhibitors}, and multi-country
SmPCs~\cite{pvverse2026}, but as human-authored comparisons, not ML benchmarks.
RegDivergence-101 is the first FDA/EMA benchmark with an explicit SILENT class.

\textbf{Foundation models and AI governance beyond text.} The momentum of
foundation models now spans multimodal scientific modelling~\cite{11541222},
vector-graphics generation~\cite{liang2026render} and animation~\cite{liang2026vanim},
and vision-language navigation~\cite{zeng2025janusvln}, while parameter-efficient
domain adaptation and ensembling of small language models~\cite{codes2026} lower
deployment cost. In high-stakes sectors such as healthcare, deep-learning
deployment is increasingly governed by policy-aligned evaluation frameworks that
weigh safety and regulatory constraints over raw accuracy~\cite{chen2026dlsg}.
RegDivergence-101 extends this momentum to cross-jurisdiction regulatory
reasoning, a high-stakes text domain that remains underrepresented in current
foundation-model evaluation.

\textbf{Gap.} Contradiction detection is intra-document or
single-jurisdiction; Graph-RAG targets single-corpus QA or
single-jurisdiction compliance; FDA/EMA divergence is documented manually but
never automated. Our task sits precisely in this gap.

%% -------------------------------------------------------------------------
\section{Task and Dataset}
%% -------------------------------------------------------------------------

\subsection{Task Definition}
\label{sec:taskdef}

For a topic $t$, let $f_t$ be an FDA requirement statement and $e_t$ the
corresponding EMA statement. The label $y_t$ is:
\begin{itemize}
\item \textbf{AGREE} --- both jurisdictions require substantively the same
  thing (same modal strength, same threshold, same scope of applicability).
  Paraphrase equivalence and synonymous modals (\textit{shall} / \textit{must})
  are treated as AGREE; minor drafting differences that do not alter sponsor
  obligations are disregarded.
\item \textbf{DIVERGE} --- the requirements directly conflict: they impose
  different thresholds, different modal strengths (\textit{shall} vs.\
  \textit{should}), different timelines, or mutually incompatible design
  specifications.
\item \textbf{SILENT} --- one jurisdiction does not address what the other
  regulates. SILENT is \textit{directional} by construction:
  \begin{itemize}
    \item \textbf{SILENT\_FDA}: FDA is silent on this topic; an EMA requirement
      exists but no corresponding FDA requirement was found in the full published
      FDA corpus (guidance documents, Q\&As, MaPPs).
    \item \textbf{SILENT\_EMA}: EMA is silent on this topic; an FDA requirement
      exists but no corresponding EMA requirement was found in the full published
      EMA corpus (guidelines, Q\&As, reflection papers).
  \end{itemize}
  RegDivergence-101 records direction in the \texttt{silent\_direction} field
  and reports per-direction F1 in Table~\ref{tab:silent_direction}. A unified
  \textbf{SILENT} label (= SILENT\_FDA $\cup$ SILENT\_EMA) is retained in
  Table~\ref{tab:results} for backward compatibility.
\end{itemize}

\textbf{Epistemic status of SILENT labels.}
A SILENT label is an absence claim: it asserts that no counterpart requirement
was found after an exhaustive corpus search. This carries irreducible epistemic
uncertainty. We bound this uncertainty through (a) a documented corpus-search
protocol (\S\ref{sec:annotation}) and (b) a per-pair \texttt{doc\_confidence}
flag (HIGH / MODERATE). Downstream work should treat MODERATE-confidence SILENT
pairs as soft labels.

\textbf{Task scope and assumptions.} This formulation assumes requirement pairs
are \textit{pre-aligned}---both the FDA and EMA sides address the same
regulatory topic. Topic alignment is a non-trivial upstream problem: FDA and
EMA guidance frequently address overlapping domains at inconsistent levels of
specificity, making it unclear whether two requirements are ``on the same
topic'' at all. RegDivergence-101 is constructed from expert-panel comparisons
that \textit{already provide aligned pairs} and from a two-track alignment
procedure described in \S\ref{sec:dataset}; we treat topic alignment as a
separate, upstream task and refer readers to \S\ref{sec:futurework} for the
corpus-level architecture that would encompass it.

\subsection{Pair Construction Protocol}
\label{sec:dataset}

RegDivergence-101 is constructed via two tracks that differ in how topical
correspondence between FDA and EMA requirements is established. Per-pair
source-track attribution (\texttt{track}: A or B) is released with the dataset.

\textbf{Track A --- expert-comparison-derived pairs (59 pairs, 58.4\%).}
All 59 pairs were extracted from three open-access peer-reviewed studies whose
explicit purpose was comparative mapping of FDA and EMA requirements in a
defined therapeutic domain:

\begin{enumerate}
\item \citet{uc2025jcc} --- ulcerative colitis endpoints and trial design
  (40 pairs).
\item \citet{pmc8157504} --- psychiatric drug-trial guidelines (10 pairs).
\item \citet{pmc6529498} --- radiopharmaceutical nonclinical requirements
  (9 pairs).
\end{enumerate}

In each source paper, domain experts had already established co-topicality
through structured literature mapping or side-by-side dossier comparison.
Our annotation task for Track A was to classify the \textit{relationship}
(AGREE / DIVERGE / SILENT) between requirements that the source authors had
already deemed co-topical. We inherit, document, and cite their alignment
judgements; per-pair source-study attribution is released with the dataset.

\textbf{Track B --- primary guidance pairs (42 pairs, 41.6\%).}
For pairs derived from ICH E9(R1), E11A~\cite{iche11a2024}, and FDA/EMA
biosimilar and gene-therapy guidance, topical alignment was established in
two steps:

\begin{enumerate}
\item \textit{Anchor identification.} Requirements were associated with ICH
  Common Technical Document (CTD) module and sub-topic designations (e.g.,
  Module 5.3.5.1 for comparative bioavailability study reports). Requirements
  sharing the same CTD anchor were candidate pairs.
\item \textit{Expert confirmation.} The primary annotator independently
  reviewed each candidate pair and confirmed or rejected co-topicality before
  the pair entered the dataset.
\end{enumerate}

\textbf{Specificity exclusion criterion.}
A candidate Track B pair was excluded if the FDA and EMA requirements operated
at different levels of the regulatory hierarchy (e.g., one requirement
addressing trial design at the \textit{statistical analysis plan} level while
the other addressed the same concept at the level of \textit{general
statistical principles}). Such pairs would conflate regulatory
level-of-detail differences with substantive DIVERGE labels.
A small number of candidate pairs were excluded on this basis (the count and
representative cases are documented in the released dataset's
\texttt{excluded\_pairs.jsonl} file). The complete exclusion-criteria checklist
is released with the dataset so third parties can audit or extend our alignment
judgements.

Each pair records
\texttt{\{id, topic, track, source, fda\_text, ema\_text, label,
silent\_direction, doc\_confidence, disputed, annotation\_notes\}}.
The combined release contains \textbf{101 pairs}: 42 AGREE, 38 DIVERGE, 21
SILENT, across 13 topic areas. We report macro-F1 to weight all classes equally.

\subsection{Annotation Protocol}
\label{sec:annotation}

\textbf{Annotators.} Annotator~A (C.~Wu) is a solution architect for a
regulatory-document AI platform; Annotator~B (Z.~Zhou) is a product manager for
an AI-powered clinical-trial management platform. Neither holds formal RA
credentialing; Track A expert-panel papers serve as the external human-expert
anchor.

\textbf{Stages.} Stage~1: Wu assigned labels using a definitions-based framing
(Track A labels inferred from source authors' stated assessments; Track B labels
assigned against full guidance text). Stage~2: Zhou independently re-labelled
all 101 pairs under a sponsor-obligation framing to probe framing-independent
stability. Both annotators are co-authors; the $\kappa$ is a cross-framing
consistency measure, not identical-instruction IAA. An independent
third-annotator pass is planned for RegDivergence-500.

\textbf{IAA.} Human cross-framing $\kappa = 0.85$ (3-class, 101 pairs;
observed agreement 0.90)---\textit{substantial} per \citet{landis1977kappa},
on par with MedNLI (0.65) and ContractNLI ($\approx$0.72). SILENT is
near-perfect ($\kappa = 0.98$); AGREE/DIVERGE more variable ($\approx$0.79--0.81).
A secondary LLM prompt-robustness check on a 36-pair sample yielded
$\kappa = 0.542$ (LLM test-retest, not human IAA).

\textbf{Disagreements (n = 10).} Six resolved by source-paper language; four
genuinely borderline pairs retain the Stage~1 label and are flagged
\texttt{disputed = True} in the dataset with written justifications.

\textbf{SILENT confidence.} All 21 SILENT pairs carry
\texttt{doc\_confidence = MODERATE}: exhaustive corpus search was performed
(full FDA and EMA guidance, Q\&As, reflection papers) but absence claims remain
epistemically soft. The HIGH/MODERATE audit is deferred to RegDivergence-500.

%% -------------------------------------------------------------------------
\section{Baselines}
%% -------------------------------------------------------------------------

\subsection{Evaluation Context}
\label{sec:eval_context}

RegDivergence-101 is an \textbf{evaluation benchmark}, not a training dataset.
All four baselines are zero-shot or few-shot. Adjacent methods are separated by
$\Delta = 0.15$--$0.60$ macro-F1 (min: Graph-RAG vs.\ LLM, $\Delta = 0.167$;
CIs partially overlapping); bootstrap CI half-widths $\approx \pm 0.08$--$0.10$
at $n = 101$. Results are reported as a ranking trend, not pairwise significance.

\subsection{Shared Evaluation Protocol}

We evaluate four methods under a shared reporting framework:
\textbf{stratified 5-fold splits} averaged over 5 random seeds, with
\textbf{bootstrap 95\% CI} (1{,}000 resamples). The protocol differs by
method class. The lexical heuristic and NLI cross-encoder have tunable
thresholds and use \textbf{nested} CV (inner folds tune thresholds, outer folds
report). The Graph-RAG classifier and LLM judge have no tunable
hyperparameters: the Graph-RAG classifier is reported over the same outer
folds, and the LLM judge is scored directly on all 101 pairs. CIs for both LLM
methods are bootstrap-only.

\subsection{Lexical Heuristic}

No neural model, no API---fully offline and deterministic. Combines TF-IDF
cosine similarity with five conflict signals: negation asymmetry, divergence
cue pairs (e.g., \textit{single-arm} vs.\ \textit{randomised}), modal-strength
asymmetry (\textit{must/shall} vs.\ \textit{may/encouraged}),
numeric-threshold mismatch, and one-sided specificity (concrete number
vs.\ case-by-case deferral). A shared ICH citation is a strong AGREE signal.

\subsection{NLI Cross-Encoder}

\texttt{typeform/distilbert-base-uncased-mnli}, run in both directions; labels
derived from contradiction/neutral/entailment probabilities with CV-tuned
thresholds. We use DistilBERT-MNLI as a representative off-the-shelf NLI
system. The structural SILENT gap observed in \S\ref{sec:discussion} (Observation~1)
is model-agnostic: any NLI model lacking an explicit SILENT output class will
map absence-of-regulation to \textit{neutral}, collapsing SILENT recall.
Substituting DeBERTa-v3-large or a modern cross-encoder would improve
AGREE/DIVERGE F1 but would not resolve the SILENT structural gap without
task-specific supervision.

\subsection{Graph-RAG Pair-Level Classifier}

For each text, Claude Haiku extracts a policy triplet
$\langle$\textit{subject, obligation\_level, requirement, conditions}$\rangle$
where $\text{obligation\_level} \in \{\text{MANDATORY}, \text{PROHIBITED},
\text{RECOMMENDED}, \text{PERMITTED}, \text{SILENT}\}$. Only
MANDATORY~$\leftrightarrow$~PROHIBITED is a hard-coded DIVERGE rule; all other
pairs are escalated to a second LLM call that reasons over the
\textit{obligation structure} of the aligned node pair, with explicit
calibration that modal-register differences (RECOMMENDED vs.\ MANDATORY) do not
automatically imply DIVERGE.

\textbf{Within-model scope.} Both the Graph-RAG classifier and the flat LLM
judge use \texttt{claude-haiku-4-5-20251001}. The comparison between them is
therefore a \textit{within-model ablation}---with vs.\ without explicit
obligation-graph structure---not an architectural comparison across model
families. Conclusions about the relative value of graph structure should be
confirmed with a second model family before generalisation.

\subsection{LLM Judge (Flat Pairwise)}

Claude Haiku with explicit AGREE / DIVERGE / SILENT definitions, prompted once
per pair with both texts. No graph, no retrieval. Full prompt and code at
Appendix~\ref{app:prompt}.

%% -------------------------------------------------------------------------
\section{Results}
\label{sec:results}
%% -------------------------------------------------------------------------

Table~\ref{tab:results} reports macro-F1, accuracy, and per-class F1 with
bootstrap 95\% confidence intervals for all four methods.

\begin{table*}
  \caption{Macro-F1, accuracy, and per-class F1 with bootstrap 95\% confidence
    intervals for the four baselines ($n = 101$ pairs).}
  \label{tab:results}
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{llllll}
    \toprule
    Method & Macro-F1 [95\% CI] & Acc & AGREE F1 [CI] & DIVERGE F1 [CI] & SILENT F1 [CI] \\
    \midrule
    Lexical heuristic$^1$
      & 0.511 [0.411--0.605] & 0.541
      & 0.718 [0.605--0.809] & 0.466 [0.320--0.593] & 0.334 [0.133--0.537] \\
    NLI cross-encoder$^2$
      & 0.233 [0.152--0.303] & 0.238
      & 0.279 [0.103--0.449] & 0.383 [0.247--0.506] & 0.025 [0.000--0.086] \\
    Graph-RAG pair-level$^3$
      & 0.663 [0.570--0.747] & 0.673
      & 0.754 [0.646--0.851] & 0.586 [0.433--0.722] & 0.640 [0.478--0.771] \\
    \textbf{LLM judge (Haiku)}$^4$
      & \textbf{0.830 [0.747--0.908]} & \textbf{0.842}
      & \textbf{0.905 [0.831--0.965]} & \textbf{0.805 [0.696--0.892]}
      & \textbf{0.778 [0.609--0.914]} \\
    \bottomrule
  \end{tabular}}
  \smallskip\\
  {\small $^1$Nested 5-fold CV, 5 seeds; thresholds tuned on inner folds only.
  $^2$Same CV protocol, thresholds tuned on inner folds.
  $^3$No tunable hyperparameters; bootstrap CI only.
  $^4$\texttt{claude-haiku-4-5-20251001}, default temperature; bootstrap CI
  only. Estimated cost: {\raise.17ex\hbox{$\scriptstyle\sim$}}\$0.10 for all
  101 pairs.}
\end{table*}

\begin{table}[h]
  \caption{Per-direction SILENT F1 for all four methods (bootstrap 95\% CI,
    $n=101$ pairs). SILENT\_FDA ($n=13$): FDA is silent, EMA regulates.
    SILENT\_EMA ($n=8$): EMA is silent, FDA regulates.}
  \label{tab:silent_direction}
  \small
  \begin{tabular}{lll}
    \toprule
    Method & SILENT\_FDA F1 [95\% CI] & SILENT\_EMA F1 [95\% CI] \\
    \midrule
    Lexical heuristic
      & 0.277 [0.000--0.545] & 0.393 [0.000--0.706] \\
    NLI cross-encoder
      & 0.049 [0.000--0.167] & 0.000 [0.000--0.000] \\
    Graph-RAG pair-level
      & 0.716 [0.516--0.882] & 0.535 [0.267--0.762] \\
    \textbf{LLM judge (Haiku)}
      & \textbf{0.795 [0.588--0.952]} & \textbf{0.738 [0.429--0.952]} \\
    \bottomrule
  \end{tabular}
  \smallskip\\
  {\small SILENT-FDA: 13 pairs (FDA silent, EMA active). SILENT-EMA: 8 pairs
    (EMA silent, FDA active). Wide CIs on SILENT-EMA reflect the small stratum
    ($n=8$). Bootstrap 1{,}000 resamples; stratified by 3-class label.}
\end{table}

Point estimates increase monotonically from lexical to Graph-RAG to the LLM
judge. At $n = 101$ the marginal bootstrap CIs partially overlap, so we report
these as a consistent ranking trend rather than a claim of statistical
significance; a paired significance test is left to future work.
\textbf{Lexical ceiling.} Characteristic errors share the signature of
\textit{high lexical overlap, opposed regulatory stance}: FDA
``modified Mayo score of 5--9'' vs.\ EMA ``full Mayo score of 9--12''
(same vocabulary, different instrument); FDA 15-year LTF vs.\ EMA
case-by-case (same topic, different specificity). Resolving these requires
representing what each requirement \textit{asserts}, not surface similarity.

%% -------------------------------------------------------------------------
\section{Analysis}
\label{sec:discussion}
%% -------------------------------------------------------------------------

\subsection{What the Baseline Hierarchy Reveals}

\textbf{Observation 1 --- SILENT is semantically detectable but invisible to
NLI framing.} SILENT F1 = 0.025 under NLI (no native SILENT class) vs.\ 0.778
under the LLM judge (explicit definition). This contrast is large and consistent
across bootstrap resamples; it reflects a structural constraint---the SILENT
class requires a formulation that can express absence, which NLI entailment
cannot---rather than a model-capacity effect. We note this observation is based
on a distilbert-class model; a higher-capacity NLI encoder would improve
AGREE/DIVERGE F1 but would not resolve the absence-detection gap without
task-specific supervision (\S\ref{sec:discussion}).

\textbf{Observation 2 --- For Claude Haiku on this dataset, pair-level
obligation graphs improve over lexical (+0.15 F1) but trail flat prompting
($-$0.17 F1); CIs partially overlap at pilot scale.} The graph makes the
obligation structure explicit (MANDATORY / RECOMMENDED / etc.), which helps
distinguish clear stance differences from similar-sounding text. But it
discards full-text context that flat prompting uses to separate
\textit{same-requirement / different-modal-register} (AGREE) from
\textit{same-topic / different-threshold} (DIVERGE). Pairs like gene-01 (FDA
15 years vs.\ EMA case-by-case) both extract as RECOMMENDED and the graph
calls them AGREE; the flat LLM correctly identifies the threshold conflict from
the text~\citep{qian2026evidenceforce}.
\textit{Scope note.} Both the graph classifier and the flat LLM judge use
Claude Haiku; this comparison is a within-model ablation (with vs.\ without
explicit graph structure) rather than a general claim about graph methods vs.\
LLM methods. Replication with a second model family is left to future work.

\textbf{Observation 3 --- Two distinct problem scopes suggest two distinct
architectural targets.} The \textit{pair-level scope} --- given pre-aligned
sentences, label AGREE/DIVERGE/SILENT --- is largely addressed by the flat
LLM judge (0.830), approaching a task ceiling. The \textit{corpus-level scope}
--- ``is EMA silent on this topic \textit{anywhere} across its guideline
corpus?'' --- requires finding aligned pairs, not just classifying them; a
pairwise LLM cannot address this, but a corpus-level policy graph can. These
are different tasks. This paper establishes baselines for the pair-level scope;
\S\ref{sec:futurework} motivates the corpus-level architecture as the next
contribution. This observation is architectural and directional; it is not a
validated finding.

\subsection{Source-Stratified Performance Analysis}
\label{sec:source_stratified}

To assess whether performance reflects source-study-specific patterns rather
than genuine task competence, we analyse performance stratified by track and
by source concentration.

\textbf{Track-level.} On AGREE/DIVERGE classes, the LLM judge scores
\textit{lower} on Track A (0.769/0.745) than Track B (0.966/0.923): exposure
to source studies does not inflate performance. The macro-F1 gap (Track B 0.630
vs.\ Track A 0.765) is explained by SILENT support ($n = 0$ vs.\ $n = 21$),
not memorisation.

\textbf{Source-concentration.} 40/101 pairs come from one source
(\citeauthor{uc2025jcc}). LLM judge on the JCC stratum ($n = 40$): 0.804;
on all other pairs ($n = 61$): 0.844. Performance is marginally
\textit{higher} outside JCC, ruling out domain-concentration inflation.

%% -------------------------------------------------------------------------
\section{Future Work}
\label{sec:futurework}
%% -------------------------------------------------------------------------

\textbf{Corpus-level policy graph.}
The correct long-term target is corpus-level silent detection and divergence
discovery: (1) build per-jurisdiction policy graphs indexing all FDA/EMA
guidance at corpus scale; (2) align requirement nodes via entity normalisation
and embedding similarity; (3) classify aligned pairs as AGREE/DIVERGE/SILENT by
obligation-structure comparison; (4) generate natural-language explanations.
This architecture subsumes the pair-alignment assumption: alignment becomes a
graph-construction step, not a pre-processing input.

\textbf{RegDivergence-500.}
Four domains are unrepresented in the pilot: CMC, adaptive trial design/estimands,
statistical analysis plan requirements, and labelling/SmPC. RegDivergence-500
($\approx$500 pairs) will cover all four via a community-annotation pipeline
seeded from a $\sim$45-source FDA/EMA comparison inventory.
Annotation guidelines are released with this paper. A third jurisdiction (PMDA)
via FDA/EMA/PMDA concordance studies~\cite{mita2026doses,gomezfernandez2026inhibitors}
will add an $\approx$40-pair FDA$\times$PMDA subset. Multi-model replication and
an independent third-annotator validation pass are planned.
A cross-pair generalisation probe against Sino-US-DrugQA~\cite{sinousdrugqa2026}
will be the first test of whether conclusions transfer beyond the FDA/EMA pair.

%% -------------------------------------------------------------------------
\section{Limitations}
\label{sec:limitations}
%% -------------------------------------------------------------------------

\textbf{Topic alignment.} Track B co-topicality judgements are single-annotator;
Track A inherits source-study alignment methodology. Reproducibility depends on
the released exclusion-criteria checklist (\texttt{data/alignment\_protocol.md}).

\textbf{Scale and coverage.} 101 pairs from three source studies limits
statistical power (directional observations, not confirmed findings). Oncology,
rare diseases, medical devices, and vaccines are absent. RegDivergence-500
(\S\ref{sec:futurework}) is required before domain-general conclusions.

\textbf{SILENT labels.} All 21 SILENT pairs carry \texttt{doc\_confidence = MODERATE};
treat them as soft labels. The pair-level SILENT task is a proxy for the correct
corpus-level formulation (\S\ref{sec:futurework}).

\textbf{Single-model evaluation.} All LLM-dependent components (triplet extraction,
graph classification, flat judge, annotation check) use Claude Haiku. The
Graph-RAG vs.\ LLM comparison is a within-model ablation, not a cross-model
comparison. The $\kappa = 0.542$ LLM consistency check reflects the same model's
label tendencies. Multi-model replication is deferred to RegDivergence-500.

\textbf{Contamination.} Haiku may have seen FDA/EMA guidance during pretraining.
The source-stratified analysis (\S\ref{sec:source_stratified}) provides evidence
against inflation; future work should evaluate on post-cutoff guidance.

\textbf{Use norms.} RegDivergence-101 is a pilot benchmark: not sufficient for
deployment decisions, not domain-generalizable without held-out evaluation, and
not a sole criterion for AI procurement in regulatory-affairs workflows.
Conclusions are specific to the FDA/EMA pair until cross-pair generalisation
probes are run.

%% -------------------------------------------------------------------------
\section{Conclusion}
%% -------------------------------------------------------------------------

We introduced cross-jurisdiction regulatory divergence detection, released
\textbf{RegDivergence-101}, a 101-pair expert-grounded \textbf{pilot}
evaluation benchmark ($\kappa = 0.85$, documented two-track pair-construction
protocol, per-direction SILENT annotation), and characterised a four-method
baseline hierarchy with bootstrap CIs. Three directional observations at pilot
scale ($n = 101$, CIs partially overlapping for adjacent methods): SILENT
requires explicit absence-aware formulation (NLI F1 0.03 $\to$ LLM F1 0.78);
for Claude Haiku on this dataset, pair-level graph structure improves over
lexical heuristics (+0.15 F1) but trails flat prompting ($-$0.17 F1); and
the two-scope distinction (pair-level vs.\ corpus-level) motivates the
corpus-level graph architecture as the next contribution. The flat LLM judge
(0.830) establishes a pair-level performance reference at pilot scale; the
corpus-level architecture proposed in \S\ref{sec:futurework} is the correct
target for large-scale silence discovery. The benchmark is a pilot: four regulatory domains (CMC,
clinical trial design, statistical analysis, labelling) remain unrepresented,
and the planned RegDivergence-500 extension with expanded dual-annotation will
establish the task at production scale.

%%
%% Acknowledgments
%%
\begin{acks}
This work was conducted independently. Benchmark pairs were derived from
publicly available FDA/EMA guidance documents and open-access peer-reviewed
comparison studies. LLM API calls used the Anthropic API (Claude Haiku,
\texttt{claude-haiku-4-5-20251001}).
\end{acks}

%%
%% Bibliography
%%
\bibliographystyle{ACM-Reference-Format}
\bibliography{xjurisdiction-divergence}

%%
%% Appendix
%%
\appendix

\section{Prompts and Code}
\label{app:prompt}

Full prompts (LLM judge system/user, Graph-RAG Stage~1 triplet-extraction,
Stage~2 graph-classification), evaluation scripts, and dataset schema are
available at the project repository:
\url{https://github.com/chuchugo/cross-jurisdiction-regulatory-divergence}

\noindent\textbf{LLM judge} (\texttt{llm\_judge.py}): system prompt defines
AGREE/DIVERGE/SILENT with explicit absence framing; user prompt pairs
\texttt{\{fda\_text\}} and \texttt{\{ema\_text\}}.
Model: \texttt{claude-haiku-4-5-20251001}, temperature 1.0, max tokens 100.

\noindent\textbf{Graph-RAG} (\texttt{graph\_rag.py}): Stage~1 extracts a
policy node \texttt{\{subject, obligation\_level, requirement, conditions\}};
Stage~2 classifies the aligned node pair. Hard rules: MANDATORY
$\leftrightarrow$ PROHIBITED $\Rightarrow$ DIVERGE (no LLM call); both
SILENT $\Rightarrow$ SILENT. Max tokens: 300.

\end{document}
