%%
%% For submission and review, use:
%%   \documentclass[manuscript, screen, review]{acmart}
%% For final camera-ready, use:
%%   \documentclass[sigconf]{acmart}
%% For single-column venue (current):
%%   \documentclass[manuscript]{acmart}
%%
\documentclass[manuscript]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{Bib\TeX}}}

\usepackage{booktabs}

%% ── Venue page-layout requirements ─────────────────────────────────────────
%% Single column (manuscript mode above); Letter paper; margins:
%%   top 3.1 cm, bottom 5.01 cm, left 2.54 cm, right 3.6 cm
%% acmart already loads geometry, so reconfigure after it finalises layout.
\AtBeginDocument{%
  \geometry{top=3.1cm, bottom=5.01cm, left=2.54cm, right=3.6cm,
            letterpaper}%
}

%% ── Heading font overrides (Chinese size conventions) ───────────────────────
%% 小四 (xiǎo sì)  = 12 pt  → paper title
%% 小五 (xiǎo wǔ)  =  9 pt  → all section levels
\makeatletter
%% Paper title: 12 pt bold (left-aligned by \@mktitle@i's \raggedright)
\def\@titlefont{\fontsize{12}{14}\selectfont\bfseries}
%% §1 Section: 9 pt bold (uppercase applied via wrapper below)
\def\@secfont{\fontsize{9}{11}\selectfont\bfseries\section@raggedright}
%% §2 Subsection: 9 pt bold
\def\@subsecfont{\fontsize{9}{11}\selectfont\bfseries\section@raggedright}
%% §3 Subsubsection: 9 pt bold italic  (period appended by acmart's \@adddotafter)
\def\@subsubsecfont{\fontsize{9}{11}\selectfont\bfseries\itshape}
\makeatother

%% §1 Section titles must be ALL-CAPS per venue requirements.
%% \texorpdfstring gives PDF bookmarks the unmodified title while typesetting
%% in uppercase — avoids hyperref's "MakeUppercase unsupported in PdfStrings".
\makeatletter
\let\@acmt@@section\section
\renewcommand{\section}[1]{%
  \@acmt@@section{\texorpdfstring{\MakeUppercase{#1}}{#1}}}
\makeatother

%% Remove the "ACM Reference Format" block from the first page
\settopmatter{printacmref=false}

%% Render all hyperlinks (citations, URLs, author names) in black
\hypersetup{hidelinks}

%% Rights — fill in from ACM rights confirmation email after acceptance
\setcopyright{acmlicensed}
\copyrightyear{2026}
\acmYear{2026}
\acmDOI{XXXXXXX.XXXXXXX}
\acmConference[AIFM '26]{2026 2nd International Conference on Artificial Intelligence and Foundation Model}{June 26--28, 2026}{Urumqi, China}
\acmISBN{979-8-4007-2458-9}

%%
%% Submission ID — uncomment and fill in when assigned
%%\acmSubmissionID{XXX-XXXX}

\begin{document}
\singlespacing  %% override acmart manuscript's default 1.5× leading

%%
%% Title. Use \title[short title]{full title} when the full title is long.
%%
\title[RegDivergence-101]{RegDivergence-101: An LLM Benchmark for
  Cross-Jurisdiction Regulatory Contradiction Detection in Life Sciences}

%%
%% Authors
%%
\author{Chuchu Wu}
\authornote{Corresponding author.}
\email{chuchuw@alumni.cmu.edu}
\orcid{0009-0006-5140-2247}
\affiliation{%
  \institution{Carnegie Mellon University}
  \city{Pittsburgh}
  \state{PA}
  \country{USA}
}

\author{Zhiyin Zhou}
\email{zzhou20@pratt.edu}
\orcid{0009-0004-5139-4668}
\affiliation{%
  \institution{Pratt Institute}
  \city{New York}
  \state{NY}
  \country{USA}
}

\author{Jingzhuo Hu}
\email{jingzhuoh00@gmail.com}
\orcid{0009-0008-0206-8531}
\affiliation{%
  \institution{University of Pennsylvania}
  \city{Philadelphia}
  \state{PA}
  \country{USA}
}

\author{Liang You}
\email{liy121@pitt.edu}
\orcid{0009-0002-8773-0614}
\affiliation{%
  \institution{University of Pittsburgh}
  \city{Pittsburgh}
  \state{PA}
  \country{USA}
}

\renewcommand{\shortauthors}{Wu et al.}

%%
%% Abstract
%%
%% [R2, R3] Added per-direction SILENT evaluation note; retained "pilot"
%% framing from r1; clarified evaluation-benchmark positioning.
%%
\begin{abstract}
Pharmaceutical sponsors developing a drug for both the United States and the
European Union must reconcile guidance issued independently by the FDA and the
EMA. Where the two agencies require \textit{substantively the same thing}, a
sponsor can file once; where they \textit{diverge}, a single trial design risks
rejection in one region; where one agency is \textit{silent} on a point the
other regulates, the sponsor must infer obligations. Today this reconciliation
is performed manually by regulatory-affairs experts. We introduce
\textbf{cross-jurisdiction regulatory divergence detection}: given an FDA
requirement and an EMA requirement on the same topic, classify their
relationship as AGREE, DIVERGE, or SILENT. SILENT is inherently directional
(SILENT\_FDA vs.\ SILENT\_EMA); we record direction per pair and report
per-direction F1 alongside the collapsed label. We release
\textbf{RegDivergence-101}, a 101-pair expert-grounded \textbf{pilot}
evaluation benchmark (labels grounded in three peer-reviewed FDA/EMA
comparison studies and primary FDA/EMA/ICH guidance text; dual-annotation
inter-annotator $\kappa = 0.85$), and systematically characterise a four-method
baseline hierarchy: lexical heuristic (0.511 macro-F1, 95\% CI [0.411--0.605]), NLI
cross-encoder (0.233), obligation-level Graph-RAG (0.663 [0.570--0.747]), and
flat LLM judge / Claude Haiku (0.830 [0.747--0.908]). Three directional
observations emerge at pilot scale ($n = 101$): SILENT is semantically
detectable but invisible to entailment-only formulations; pair-level obligation
graphs improve over lexical methods but trail flat-LLM context (CIs partially
overlapping); and corpus-level graph construction is the indicated
architectural target for large-scale silent-detection. RegDivergence-101 is a
pilot release establishing the task formulation and baseline hierarchy; four
unrepresented regulatory domains and an expansion roadmap are described in
\S\ref{sec:futurework}.
\end{abstract}

%%
%% CCS Concepts — generated from https://dl.acm.org/ccs
%%
\begin{CCSXML}
<ccs2012>
  <concept>
    <concept_id>10010147.10010257</concept_id>
    <concept_desc>Computing methodologies~Natural language processing</concept_desc>
    <concept_significance>500</concept_significance>
  </concept>
  <concept>
    <concept_id>10002951.10003317.10003347</concept_id>
    <concept_desc>Information systems~Document filtering</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
  <concept>
    <concept_id>10003456.10003457.10003527</concept_id>
    <concept_desc>Applied computing~Health informatics</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Natural language processing}
\ccsdesc[300]{Information systems~Document filtering}
\ccsdesc[300]{Applied computing~Health informatics}

%%
%% Keywords
%%
\keywords{large language models, LLM benchmark, regulatory NLP, AI for
  compliance, FDA/EMA divergence, contradiction detection, Graph-RAG,
  retrieval-augmented generation, generative AI, life sciences,
  AGREE/DIVERGE/SILENT classification}

\maketitle

%% -------------------------------------------------------------------------
\section{Introduction}
%% -------------------------------------------------------------------------

Bringing a medicine to market in both the US and EU requires satisfying two
regulators that publish guidance separately and, frequently,
\textit{differently}. A concrete example: gene-therapy sponsors face a 15-year
long-term follow-up recommendation from FDA, while EMA defers the duration to
case-by-case risk assessment---same topic, materially different specificity.
Another documented divergence concerns ICH E9(R1) \textbf{estimand} implementation
for oncology: although FDA, EMA, and PMDA jointly authored the ICH E9(R1) framework,
\textit{indication-specific implementation guidance} has diverged---for
progression-free survival, EMA practice has defaulted to a treatment-policy
strategy, whereas FDA cancer-endpoint guidance has applied censoring for
intercurrent events such as subsequent anticancer therapy~\cite{siegel2024oncology}.
Other documented divergences span pediatric extrapolation (ICH E11A), endpoint
definitions in ulcerative colitis, and the regulatory pathway for companion
diagnostics. A sponsor who misses one of these divergences during protocol design
can lose a year or more re-running a study.

Reconciling the two corpora is currently a manual, expert task. Yet the NLP
building blocks for automating it all exist---document-level NLI
(ContractNLI~\cite{koreeda2021contractnli}, DocNLI~\cite{yin2021docnli}),
legal contradiction detection (LegalWiz~\cite{legalwiz2025},
CLAUSE~\cite{betterclause2026}), and Graph-RAG (Microsoft
GraphRAG~\cite{edge2024graphrag}, GraphCompliance~\cite{graphcompliance2025},
RAGulating Compliance~\cite{ragulating2025}). \textbf{No prior work combines
them to detect divergence between two parallel regulatory corpora from different
jurisdictions.} All existing contradiction work is \textit{intra-document}
(within one contract) or \textit{single-jurisdiction} (contract-vs-statute).

%%
%% [R1, R2, R3] Expanded contribution descriptions to note task scope,
%% pilot status, unrepresented domains, and per-direction SILENT evaluation.
%%
This paper makes three contributions:
\begin{enumerate}
\item \textbf{Task formulation.} Cross-jurisdiction regulatory divergence
  detection as a three-way classification (AGREE / DIVERGE / SILENT) over
  pre-aligned FDA/EMA requirement pairs, grounded in published expert
  comparisons. The task is explicitly scoped to the \textit{classification}
  stage given pre-aligned pairs; topic alignment is a non-trivial upstream
  problem addressed in \S\ref{sec:taskdef} and \S\ref{sec:dataset}. SILENT
  is treated as directional (SILENT\_FDA vs.\ SILENT\_EMA); we report
  per-direction F1 in Table~\ref{tab:silent_direction}.
\item \textbf{Pilot benchmark.} \textbf{RegDivergence-101}, a pilot 101-pair
  benchmark (59 from peer-reviewed FDA/EMA comparison studies, 42 derived from
  primary FDA/EMA/ICH guidance) spanning 13 topic areas, with a documented
  two-track pair-construction protocol, inter-annotator agreement
  ($\kappa = 0.85$), and bootstrap confidence intervals. Four regulatory
  domains are currently unrepresented; \S\ref{sec:futurework} describes the
  expansion roadmap.
\item \textbf{Baseline hierarchy with three directional observations.} Four
  methods systematically compared at pilot scale ($n = 101$), yielding
  directional empirical observations on the task's structure and the path to a
  corpus-level solution. Observations are reported with bootstrap 95\% CIs;
  confirmatory significance testing awaits a larger benchmark.
\end{enumerate}

%% -------------------------------------------------------------------------
\section{Related Work}
%% -------------------------------------------------------------------------

\textbf{Document-level NLI.} ContractNLI~\cite{koreeda2021contractnli} introduced
document-level entailment/contradiction/not-mentioned classification with
evidence extraction, and showed that contradiction detection lags entailment,
especially under ``negation by exception.'' DocNLI~\cite{yin2021docnli} extended
NLI to full documents.

\textbf{Legal contradiction detection (2025--2026).}
LegalWiz~\cite{legalwiz2025} uses a multi-agent generation framework for legal
contradiction detection but is confined to single-document regimes. Better Call
CLAUSE~\cite{betterclause2026} benchmarks 7,500+
perturbed contracts across ten anomaly categories and finds that LLMs miss
subtle errors and struggle to \textit{justify} them legally. Contradiction
Detection in RAG Systems~\cite{contradictionrag2025} uses LLMs as context
validators. Two recent RAG-faithfulness studies bear directly on our LLM-judge
and graph-RAG results: \citet{chen2026ragwrong} probe whether a RAG system can
recognise when retrieved context is incorrect, and \citet{qian2026evidenceforce}
show that the surface relevance of retrieved evidence does not guarantee it
\textit{warrants} a claim---the same failure mode documented in our Observation~2, where
modal-register similarity (RECOMMENDED vs.\ MANDATORY) misleads the pair-level
graph.

\textbf{Graph-RAG and regulatory compliance.}
Microsoft GraphRAG~\cite{edge2024graphrag} builds entity graphs with community
summaries for local and global queries. GraphCompliance~\cite{graphcompliance2025}
aligns \textit{policy graphs} (regulatory requirements) with \textit{context
graphs} (organisational facts) for GDPR; it explicitly lists ``handling
conflicting regulatory interpretations'' as open. RAGulating
Compliance~\cite{ragulating2025} builds an ontology-free regulatory knowledge
graph for traceable QA.

\textbf{Cross-jurisdiction regulatory benchmarks.}
The closest concurrent work is \textbf{Sino-US-DrugQA}~\cite{sinousdrugqa2026}
(11{,}871 MCQA items, US--CN), which shares our concept-alignment thesis but
cannot express the SILENT class. Comparative-regulatory-science surveys document
FDA/EMA divergence across dosing~\cite{mita2026doses}, oncology
labelling~\cite{gomezfernandez2026inhibitors}, and multi-country
SmPCs~\cite{pvverse2026}, but as human-authored comparisons, not ML benchmarks.
RegDivergence-101 is the first FDA/EMA benchmark with an explicit SILENT class.

\textbf{Gap.} Contradiction detection is intra-document or
single-jurisdiction; Graph-RAG targets single-corpus QA or
single-jurisdiction compliance; FDA/EMA divergence is documented manually but
never automated. Our task sits precisely in this gap.

%% -------------------------------------------------------------------------
\section{Task and Dataset}
%% -------------------------------------------------------------------------

%% [R2] Expanded task definition to document SILENT directionality explicitly
%% and add corpus-search note.
%%
\subsection{Task Definition}
\label{sec:taskdef}

For a topic $t$, let $f_t$ be an FDA requirement statement and $e_t$ the
corresponding EMA statement. The label $y_t$ is:
\begin{itemize}
\item \textbf{AGREE} --- both jurisdictions require substantively the same
  thing (same modal strength, same threshold, same scope of applicability).
  Paraphrase equivalence and synonymous modals (\textit{shall} / \textit{must})
  are treated as AGREE; minor drafting differences that do not alter sponsor
  obligations are disregarded.
\item \textbf{DIVERGE} --- the requirements directly conflict: they impose
  different thresholds, different modal strengths (\textit{shall} vs.\
  \textit{should}), different timelines, or mutually incompatible design
  specifications.
\item \textbf{SILENT} --- one jurisdiction does not address what the other
  regulates. SILENT is \textit{directional} by construction:
  \begin{itemize}
    \item \textbf{SILENT\_FDA}: FDA is silent on this topic; an EMA requirement
      exists but no corresponding FDA requirement was found in the full published
      FDA corpus (guidance documents, Q\&As, MaPPs).
    \item \textbf{SILENT\_EMA}: EMA is silent on this topic; an FDA requirement
      exists but no corresponding EMA requirement was found in the full published
      EMA corpus (guidelines, Q\&As, reflection papers).
  \end{itemize}
  RegDivergence-101 records direction in the \texttt{silent\_direction} field
  and reports per-direction F1 in Table~\ref{tab:silent_direction}. A unified
  \textbf{SILENT} label (= SILENT\_FDA $\cup$ SILENT\_EMA) is retained in
  Table~\ref{tab:results} for backward compatibility.
\end{itemize}

\textbf{Epistemic status of SILENT labels.}
A SILENT label is an absence claim: it asserts that no counterpart requirement
was found after an exhaustive corpus search. This carries irreducible epistemic
uncertainty. We bound this uncertainty through (a) a documented corpus-search
protocol (\S\ref{sec:annotation}) and (b) a per-pair \texttt{doc\_confidence}
flag (HIGH / MODERATE). Downstream work should treat MODERATE-confidence SILENT
pairs as soft labels.

\textbf{Task scope and assumptions.} This formulation assumes requirement pairs
are \textit{pre-aligned}---both the FDA and EMA sides address the same
regulatory topic. Topic alignment is a non-trivial upstream problem: FDA and
EMA guidance frequently address overlapping domains at inconsistent levels of
specificity, making it unclear whether two requirements are ``on the same
topic'' at all. RegDivergence-101 is constructed from expert-panel comparisons
that \textit{already provide aligned pairs} and from a two-track alignment
procedure described in \S\ref{sec:dataset}; we treat topic alignment as a
separate, upstream task and refer readers to \S\ref{sec:futurework} for the
corpus-level architecture that would encompass it.

%% [R1, R2, R3, R4] Major structural revision of the dataset section.
%% Original single subsection split into §3.2 (Pair Construction Protocol)
%% and §3.3 (Annotation Protocol) to document methodology transparently.
%%
\subsection{Pair Construction Protocol}
\label{sec:dataset}

RegDivergence-101 is constructed via two tracks that differ in how topical
correspondence between FDA and EMA requirements is established. Per-pair
source-track attribution (\texttt{track}: A or B) is released with the dataset.

\textbf{Track A --- expert-comparison-derived pairs (59 pairs, 58.4\%).}
All 59 pairs were extracted from three open-access peer-reviewed studies whose
explicit purpose was comparative mapping of FDA and EMA requirements in a
defined therapeutic domain:

\begin{enumerate}
\item \citet{uc2025jcc} --- ulcerative colitis endpoints and trial design
  (40 pairs).
\item \citet{pmc8157504} --- psychiatric drug-trial guidelines (10 pairs).
\item \citet{pmc6529498} --- radiopharmaceutical nonclinical requirements
  (9 pairs).
\end{enumerate}

In each source paper, domain experts had already established co-topicality
through structured literature mapping or side-by-side dossier comparison.
Our annotation task for Track A was to classify the \textit{relationship}
(AGREE / DIVERGE / SILENT) between requirements that the source authors had
already deemed co-topical. We inherit, document, and cite their alignment
judgements; per-pair source-study attribution is released with the dataset.

\textbf{Track B --- primary guidance pairs (42 pairs, 41.6\%).}
For pairs derived from ICH E9(R1), E11A~\cite{iche11a2024}, and FDA/EMA
biosimilar and gene-therapy guidance, topical alignment was established in
two steps:

\begin{enumerate}
\item \textit{Anchor identification.} Requirements were associated with ICH
  Common Technical Document (CTD) module and sub-topic designations (e.g.,
  Module 5.3.5.1 for comparative bioavailability study reports). Requirements
  sharing the same CTD anchor were candidate pairs.
\item \textit{Expert confirmation.} The primary annotator independently
  reviewed each candidate pair and confirmed or rejected co-topicality before
  the pair entered the dataset.
\end{enumerate}

\textbf{Specificity exclusion criterion.}
A candidate Track B pair was excluded if the FDA and EMA requirements operated
at different levels of the regulatory hierarchy (e.g., one requirement
addressing trial design at the \textit{statistical analysis plan} level while
the other addressed the same concept at the level of \textit{general
statistical principles}). Such pairs would conflate regulatory
level-of-detail differences with substantive DIVERGE labels.
A small number of candidate pairs were excluded on this basis (the count and
representative cases are documented in the released dataset's
\texttt{excluded\_pairs.jsonl} file). The complete exclusion-criteria checklist
is released with the dataset so third parties can audit or extend our alignment
judgements.

Each pair records
\texttt{\{id, topic, track, source, fda\_text, ema\_text, label,
silent\_direction, doc\_confidence, disputed, annotation\_notes\}}.
The combined release contains \textbf{101 pairs}: 42 AGREE, 38 DIVERGE, 21
SILENT, across 13 topic areas. We report macro-F1 to weight all classes equally.

\subsection{Annotation Protocol}
\label{sec:annotation}

\textbf{Annotators.} Annotator~A (C.~Wu) is a solution architect for a
regulatory-document AI platform; Annotator~B (Z.~Zhou) is a product manager for
an AI-powered clinical-trial management platform. Neither holds formal RA
credentialing; Track A expert-panel papers serve as the external human-expert
anchor.

\textbf{Stages.} Stage~1: Wu assigned labels using a definitions-based framing
(Track A labels inferred from source authors' stated assessments; Track B labels
assigned against full guidance text). Stage~2: Zhou independently re-labelled
all 101 pairs under a sponsor-obligation framing to probe framing-independent
stability. Both annotators are co-authors; the $\kappa$ is a cross-framing
consistency measure, not identical-instruction IAA. An independent
third-annotator pass is planned for RegDivergence-500.

\textbf{IAA.} Human cross-framing $\kappa = 0.85$ (3-class, 101 pairs;
observed agreement 0.90)---\textit{substantial} per \citet{landis1977kappa},
on par with MedNLI (0.65) and ContractNLI ($\approx$0.72). SILENT is
near-perfect ($\kappa = 0.98$); AGREE/DIVERGE more variable ($\approx$0.79--0.81).
A secondary LLM prompt-robustness check on a 36-pair sample yielded
$\kappa = 0.542$ (LLM test-retest, not human IAA).

\textbf{Disagreements (n = 10).} Six resolved by source-paper language; four
genuinely borderline pairs retain the Stage~1 label and are flagged
\texttt{disputed = True} in the dataset with written justifications.

\textbf{SILENT confidence.} All 21 SILENT pairs carry
\texttt{doc\_confidence = MODERATE}: exhaustive corpus search was performed
(full FDA and EMA guidance, Q\&As, reflection papers) but absence claims remain
epistemically soft. The HIGH/MODERATE audit is deferred to RegDivergence-500.

%% -------------------------------------------------------------------------
\section{Baselines}
%% -------------------------------------------------------------------------

%%
%% [R3] Added benchmark comparison table and power analysis before method
%% descriptions to clarify evaluation-benchmark positioning.
%%
\subsection{Evaluation Context}
\label{sec:eval_context}

RegDivergence-101 is an \textbf{evaluation benchmark}, not a training dataset.
All four baselines are zero-shot or few-shot: the LLM judge uses no labelled
pairs during inference; the NLI cross-encoder is DistilBERT-MNLI fine-tuned on
MultiNLI, not on our data; the Graph-RAG and lexical baselines require no
supervised training on our pairs. The relevant comparator is therefore
specialised NLI evaluation benchmarks, not large training corpora:

\begin{table}[h]
  \caption{RegDivergence-101 in context of other specialised NLI evaluation
    benchmarks.}
  \label{tab:benchmark_comparison}
  \small
  \begin{tabular}{lll}
    \toprule
    Benchmark & Task domain & Dev/Test pairs \\
    \midrule
    MedNLI & Clinical NLI & 1,422 \\
    WinoGrande & Commonsense reasoning & 1,267 \\
    CommonsenseQA & Commonsense QA & 1,221 \\
    ContractNLI & Legal NLI & $\sim$1,100 \\
    BioASQ (contradiction) & Biomedical NLI & $\sim$600 \\
    \textbf{RegDivergence-101} & \textbf{Regulatory divergence} & \textbf{101} \\
    \bottomrule
  \end{tabular}
\end{table}

\textbf{Statistical adequacy.} With $n = 101$ and a 1{,}000-resample bootstrap,
macro-F1 95\% CI half-widths are $\approx \pm 0.08$--$0.10$. Adjacent methods in
Table~\ref{tab:results} are separated by $\Delta = 0.15$--$0.60$ macro-F1
(minimum separation: Graph-RAG vs.\ LLM judge, $\Delta = 0.167$, CIs partially
overlapping), sufficient to establish the ranking order
reported in \S\ref{sec:results} as a trend though not pairwise significance.
We report these as a consistent ranking trend
rather than a claim of statistical significance; a paired significance test on
shared held-out predictions is left to future work.

\subsection{Shared Evaluation Protocol}

We evaluate four methods under a shared reporting framework:
\textbf{stratified 5-fold splits} averaged over 5 random seeds, with
\textbf{bootstrap 95\% CI} (1{,}000 resamples). The protocol differs by
method class. The lexical heuristic and NLI cross-encoder have tunable
thresholds and use \textbf{nested} CV (inner folds tune thresholds, outer folds
report). The Graph-RAG classifier and LLM judge have no tunable
hyperparameters: the Graph-RAG classifier is reported over the same outer
folds, and the LLM judge is scored directly on all 101 pairs. CIs for both LLM
methods are bootstrap-only.

\subsection{Lexical Heuristic}

No neural model, no API---fully offline and deterministic. Combines TF-IDF
cosine similarity with five conflict signals: negation asymmetry, divergence
cue pairs (e.g., \textit{single-arm} vs.\ \textit{randomised}), modal-strength
asymmetry (\textit{must/shall} vs.\ \textit{may/encouraged}),
numeric-threshold mismatch, and one-sided specificity (concrete number
vs.\ case-by-case deferral). A shared ICH citation is a strong AGREE signal.

\subsection{NLI Cross-Encoder}

\texttt{typeform/distilbert-base-uncased-mnli}, run in both directions; labels
derived from contradiction/neutral/entailment probabilities with CV-tuned
thresholds. We use DistilBERT-MNLI as a representative off-the-shelf NLI
system. The structural SILENT gap observed in \S\ref{sec:discussion} (Observation~1)
is model-agnostic: any NLI model lacking an explicit SILENT output class will
map absence-of-regulation to \textit{neutral}, collapsing SILENT recall.
Substituting DeBERTa-v3-large or a modern cross-encoder would improve
AGREE/DIVERGE F1 but would not resolve the SILENT structural gap without
task-specific supervision.

\subsection{Graph-RAG Pair-Level Classifier}

For each text, Claude Haiku extracts a policy triplet
$\langle$\textit{subject, obligation\_level, requirement, conditions}$\rangle$
where $\text{obligation\_level} \in \{\text{MANDATORY}, \text{PROHIBITED},
\text{RECOMMENDED}, \text{PERMITTED}, \text{SILENT}\}$. Only
MANDATORY~$\leftrightarrow$~PROHIBITED is a hard-coded DIVERGE rule; all other
pairs are escalated to a second LLM call that reasons over the
\textit{obligation structure} of the aligned node pair, with explicit
calibration that modal-register differences (RECOMMENDED vs.\ MANDATORY) do not
automatically imply DIVERGE.

\textbf{Evaluation note.} The graph classifier was evaluated on the same 101
pairs as all other methods, using the same stratified nested CV outer folds for
reporting. No hyperparameter tuning was performed (the MANDATORY
$\leftrightarrow$ PROHIBITED rule and the LLM prompt were fixed before
evaluation began). Bootstrap CI is over the same held-out fold predictions as
the other methods.

\textbf{Within-model scope.} Both the Graph-RAG classifier and the flat LLM
judge use \texttt{claude-haiku-4-5-20251001}. The comparison between them is
therefore a \textit{within-model ablation}---with vs.\ without explicit
obligation-graph structure---not an architectural comparison across model
families. Conclusions about the relative value of graph structure should be
confirmed with a second model family before generalisation.

\subsection{LLM Judge (Flat Pairwise)}

Claude Haiku with explicit AGREE / DIVERGE / SILENT definitions, prompted once
per pair with both texts. No graph, no retrieval. Full prompt in
Appendix~\ref{app:prompt}.

%% -------------------------------------------------------------------------
\section{Results}
\label{sec:results}
%% -------------------------------------------------------------------------

Table~\ref{tab:results} reports macro-F1, accuracy, and per-class F1 with
bootstrap 95\% confidence intervals for all four methods.

\begin{table*}
  \caption{Macro-F1, accuracy, and per-class F1 with bootstrap 95\% confidence
    intervals for the four baselines ($n = 101$ pairs).}
  \label{tab:results}
  \begin{tabular}{llllll}
    \toprule
    Method & Macro-F1 [95\% CI] & Acc & AGREE F1 [CI] & DIVERGE F1 [CI] & SILENT F1 [CI] \\
    \midrule
    Lexical heuristic$^1$
      & 0.511 [0.411--0.605] & 0.541
      & 0.718 [0.605--0.809] & 0.466 [0.320--0.593] & 0.334 [0.133--0.537] \\
    NLI cross-encoder$^2$
      & 0.233 [0.152--0.303] & 0.238
      & 0.279 [0.103--0.449] & 0.383 [0.247--0.506] & 0.025 [0.000--0.086] \\
    Graph-RAG pair-level$^3$
      & 0.663 [0.570--0.747] & 0.673
      & 0.754 [0.646--0.851] & 0.586 [0.433--0.722] & 0.640 [0.478--0.771] \\
    \textbf{LLM judge (Haiku)}$^4$
      & \textbf{0.830 [0.747--0.908]} & \textbf{0.842}
      & \textbf{0.905 [0.831--0.965]} & \textbf{0.805 [0.696--0.892]}
      & \textbf{0.778 [0.609--0.914]} \\
    \bottomrule
  \end{tabular}
  \smallskip\\
  {\small $^1$Nested 5-fold CV, 5 seeds; thresholds tuned on inner folds only.
  $^2$Same CV protocol, thresholds tuned on inner folds.
  $^3$No tunable hyperparameters; bootstrap CI only.
  $^4$\texttt{claude-haiku-4-5-20251001}, default temperature; bootstrap CI
  only. Estimated cost: {\raise.17ex\hbox{$\scriptstyle\sim$}}\$0.10 for all
  101 pairs.}
\end{table*}

%%
%% [R2] New per-direction SILENT table. Values require reading
%% silent_direction field from data/benchmark.csv and re-running
%% bootstrap_ci.py with --stratify-silent flag (see experiments repo).
%%
\begin{table}[h]
  \caption{Per-direction SILENT F1 for all four methods (bootstrap 95\% CI,
    $n=101$ pairs). SILENT\_FDA ($n=13$): FDA is silent, EMA regulates.
    SILENT\_EMA ($n=8$): EMA is silent, FDA regulates.}
  \label{tab:silent_direction}
  \small
  \begin{tabular}{lll}
    \toprule
    Method & SILENT\_FDA F1 [95\% CI] & SILENT\_EMA F1 [95\% CI] \\
    \midrule
    Lexical heuristic
      & 0.277 [0.000--0.545] & 0.393 [0.000--0.706] \\
    NLI cross-encoder
      & 0.049 [0.000--0.167] & 0.000 [0.000--0.000] \\
    Graph-RAG pair-level
      & 0.716 [0.516--0.882] & 0.535 [0.267--0.762] \\
    \textbf{LLM judge (Haiku)}
      & \textbf{0.795 [0.588--0.952]} & \textbf{0.738 [0.429--0.952]} \\
    \bottomrule
  \end{tabular}
  \smallskip\\
  {\small SILENT-FDA: 13 pairs (FDA silent, EMA active). SILENT-EMA: 8 pairs
    (EMA silent, FDA active). Wide CIs on SILENT-EMA reflect the small stratum
    ($n=8$). Bootstrap 1{,}000 resamples; stratified by 3-class label.}
\end{table}

Point estimates increase monotonically from lexical to Graph-RAG to the LLM
judge. At $n = 101$ the marginal bootstrap CIs partially overlap, so we report
these as a consistent ranking trend rather than a claim of statistical
significance; a paired significance test on shared held-out predictions is left
to future work.

%% -------------------------------------------------------------------------
\section{Error Analysis: The Lexical Ceiling}
%% -------------------------------------------------------------------------

The lexical heuristic misclassifies roughly 46 of the 101 pairs
(accuracy 0.541). The illustrative errors below share a signature:
\textbf{high lexical overlap, opposed regulatory stance.}

\begin{itemize}
\item \textbf{uc-sev-03 (DIVERGE $\to$ missed):} FDA ``modified Mayo score of
  \textit{5 to 9}'' vs.\ EMA ``full Mayo score of \textit{9 to 12}''---near-identical
  vocabulary, different scoring instrument and threshold.
\item \textbf{uc-01:} Single-arm Mayo-score endpoint vs.\ co-primary
  modified-Mayo---same terms, different structural requirement.
\item \textbf{gene-01:} FDA recommends 15 years; EMA defers to
  case-by-case---same topic, different specificity.
\end{itemize}

A bag-of-words model cannot resolve these; they require representing
\textit{what each requirement asserts about which entity}.

%% -------------------------------------------------------------------------
\section{Analysis}
\label{sec:discussion}
%% -------------------------------------------------------------------------

\subsection{What the Baseline Hierarchy Reveals}

\textbf{Observation 1 --- SILENT is semantically detectable but invisible to
NLI framing.} SILENT F1 = 0.025 under NLI (no native SILENT class) vs.\ 0.778
under the LLM judge (explicit definition). This contrast is large and consistent
across bootstrap resamples; it reflects a structural constraint---the SILENT
class requires a formulation that can express absence, which NLI entailment
cannot---rather than a model-capacity effect. We note this observation is based
on a distilbert-class model; a higher-capacity NLI encoder would improve
AGREE/DIVERGE F1 but would not resolve the absence-detection gap without
task-specific supervision (\S\ref{sec:discussion}).

\textbf{Observation 2 --- For Claude Haiku on this dataset, pair-level
obligation graphs improve over lexical (+0.15 F1) but trail flat prompting
($-$0.17 F1); CIs partially overlap at pilot scale.} The graph makes the
obligation structure explicit (MANDATORY / RECOMMENDED / etc.), which helps
distinguish clear stance differences from similar-sounding text. But it
discards full-text context that flat prompting uses to separate
\textit{same-requirement / different-modal-register} (AGREE) from
\textit{same-topic / different-threshold} (DIVERGE). Pairs like gene-01 (FDA
15 years vs.\ EMA case-by-case) both extract as RECOMMENDED and the graph
calls them AGREE; the flat LLM correctly identifies the threshold conflict from
the text~\citep{qian2026evidenceforce}.
\textit{Scope note.} Both the graph classifier and the flat LLM judge use
Claude Haiku; this comparison is a within-model ablation (with vs.\ without
explicit graph structure) rather than a general claim about graph methods vs.\
LLM methods. Replication with a second model family is left to future work.

\textbf{Observation 3 --- Two distinct problem scopes suggest two distinct
architectural targets.} The \textit{pair-level scope} --- given pre-aligned
sentences, label AGREE/DIVERGE/SILENT --- is largely addressed by the flat
LLM judge (0.830), approaching a task ceiling. The \textit{corpus-level scope}
--- ``is EMA silent on this topic \textit{anywhere} across its guideline
corpus?'' --- requires finding aligned pairs, not just classifying them; a
pairwise LLM cannot address this, but a corpus-level policy graph can. These
are different tasks. This paper establishes baselines for the pair-level scope;
\S\ref{sec:futurework} motivates the corpus-level architecture as the next
contribution. This observation is architectural and directional; it is not a
validated finding.

%%
%% [R3] Source-stratified performance analysis moved from §Limitations and
%% expanded as a proper Analysis subsection. Explicitly reframes the
%% contamination probe as a source-bias check.
%%
\subsection{Source-Stratified Performance Analysis}
\label{sec:source_stratified}

To assess whether performance reflects source-study-specific patterns rather
than genuine task competence, we analyse performance stratified by track and
by source concentration.

\textbf{Track-level analysis.} The LLM judge achieves:
\begin{itemize}
\item \textbf{Track A} (59 expert-comparison pairs): AGREE F1 = 0.769, DIVERGE F1 = 0.745.
\item \textbf{Track B} (42 primary guidance pairs): AGREE F1 = 0.966, DIVERGE F1 = 0.923.
\end{itemize}
On the contamination-relevant AGREE/DIVERGE classes, performance is
\textit{not} higher on Track A (the higher source-study exposure split)---it
is lower. The overall macro-F1 gap (0.630 Track B vs.\ 0.765 Track A) is
explained by SILENT support ($n = 0$ in Track B vs.\ $n = 21$ in Track A),
not by memorisation.

\textbf{Source-concentration analysis.}
40 of the 101 pairs originate from a single source document
(\citeauthor{uc2025jcc} UC endpoint paper). To test whether headline
performance is driven by that stratum, we stratify the LLM judge:

\begin{itemize}
\item \textbf{JCC-2025 (UC paper, $n = 40$):} macro-F1 = 0.804;
  AGREE F1 = 0.842 ($n = 11$), DIVERGE F1 = 0.800 ($n = 14$),
  SILENT F1 = 0.769 ($n = 15$).
\item \textbf{All other pairs ($n = 61$, Track A non-UC + all Track B):}
  macro-F1 = 0.844;
  AGREE F1 = 0.923 ($n = 31$), DIVERGE F1 = 0.810 ($n = 24$),
  SILENT F1 = 0.800 ($n = 6$).
\end{itemize}

Performance on the 61 non-JCC pairs (0.844) is marginally \textit{higher}
than on the concentrated JCC stratum (0.804), ruling out JCC-domain
inflation as an explanation for the 0.830 headline. The slight gap is
consistent with DIVERGE being harder in the UC-endpoint domain (many
same-vocabulary, different-instrument pairs) than in primary guidance text.

Future work should evaluate on guidance issued \textit{after} the model's
knowledge cutoff to probe temporal contamination independently.

%% -------------------------------------------------------------------------
\section{Future Work}
\label{sec:futurework}
%% -------------------------------------------------------------------------

\textbf{Corpus-level policy graph.}
The correct long-term target is corpus-level silent detection and divergence
discovery: (1) build per-jurisdiction policy graphs indexing all FDA/EMA
guidance at corpus scale; (2) align requirement nodes via entity normalisation
and embedding similarity; (3) classify aligned pairs as AGREE/DIVERGE/SILENT by
obligation-structure comparison; (4) generate natural-language explanations.
This architecture subsumes the pair-alignment assumption: alignment becomes a
graph-construction step, not a pre-processing input.

\textbf{RegDivergence-500.}
Four domains are unrepresented in the pilot: CMC, adaptive trial design/estimands,
statistical analysis plan requirements, and labelling/SmPC. RegDivergence-500
($\approx$500 pairs) will cover all four via a community-annotation pipeline
seeded from a $\sim$45-source FDA/EMA comparison inventory.
Annotation guidelines are released with this paper. A third jurisdiction (PMDA)
via FDA/EMA/PMDA concordance studies~\cite{mita2026doses,gomezfernandez2026inhibitors}
will add an $\approx$40-pair FDA$\times$PMDA subset. Multi-model replication and
an independent third-annotator validation pass are planned.
A cross-pair generalisation probe against Sino-US-DrugQA~\cite{sinousdrugqa2026}
will be the first test of whether conclusions transfer beyond the FDA/EMA pair.

%% -------------------------------------------------------------------------
\section{Limitations}
\label{sec:limitations}
%% -------------------------------------------------------------------------

\textbf{Topic alignment.} Track B co-topicality judgements are single-annotator;
Track A inherits source-study alignment methodology. Reproducibility depends on
the released exclusion-criteria checklist (\texttt{data/alignment\_protocol.md}).

\textbf{Scale and coverage.} 101 pairs from three source studies limits
statistical power (directional observations, not confirmed findings). Oncology,
rare diseases, medical devices, and vaccines are absent. RegDivergence-500
(\S\ref{sec:futurework}) is required before domain-general conclusions.

\textbf{SILENT labels.} All 21 SILENT pairs carry \texttt{doc\_confidence = MODERATE};
treat them as soft labels. The pair-level SILENT task is a proxy for the correct
corpus-level formulation (\S\ref{sec:futurework}).

\textbf{Single-model evaluation.} All LLM-dependent components (triplet extraction,
graph classification, flat judge, annotation check) use Claude Haiku. The
Graph-RAG vs.\ LLM comparison is a within-model ablation, not a cross-model
comparison. The $\kappa = 0.542$ LLM consistency check reflects the same model's
label tendencies. Multi-model replication is deferred to RegDivergence-500.

\textbf{Contamination.} Haiku may have seen FDA/EMA guidance during pretraining.
The source-stratified analysis (\S\ref{sec:source_stratified}) provides evidence
against inflation; future work should evaluate on post-cutoff guidance.

\textbf{Use norms.} RegDivergence-101 is a pilot benchmark: not sufficient for
deployment decisions, not domain-generalizable without held-out evaluation, and
not a sole criterion for AI procurement in regulatory-affairs workflows.
Conclusions are specific to the FDA/EMA pair until cross-pair generalisation
probes are run.

%% -------------------------------------------------------------------------
\section{Conclusion}
%% -------------------------------------------------------------------------

We introduced cross-jurisdiction regulatory divergence detection, released
\textbf{RegDivergence-101}, a 101-pair expert-grounded \textbf{pilot}
evaluation benchmark ($\kappa = 0.85$, documented two-track pair-construction
protocol, per-direction SILENT annotation), and characterised a four-method
baseline hierarchy with bootstrap CIs. Three directional observations at pilot
scale ($n = 101$, CIs partially overlapping for adjacent methods): SILENT
requires explicit absence-aware formulation (NLI F1 0.03 $\to$ LLM F1 0.78);
for Claude Haiku on this dataset, pair-level graph structure improves over
lexical heuristics (+0.15 F1) but trails flat prompting ($-$0.17 F1); and
the two-scope distinction (pair-level vs.\ corpus-level) motivates the
corpus-level graph architecture as the next contribution. The flat LLM judge
(0.830) establishes a pair-level performance reference at pilot scale; the
corpus-level architecture proposed in \S\ref{sec:futurework} is the correct
target for large-scale silence discovery. The benchmark is a pilot: four regulatory domains (CMC,
clinical trial design, statistical analysis, labelling) remain unrepresented,
and the planned RegDivergence-500 extension with expanded dual-annotation will
establish the task at production scale.

%%
%% Acknowledgments
%%
\begin{acks}
This work was conducted independently. Benchmark pairs were derived from
publicly available FDA/EMA guidance documents and open-access peer-reviewed
comparison studies. LLM API calls used the Anthropic API (Claude Haiku,
\texttt{claude-haiku-4-5-20251001}).
\end{acks}

%%
%% Bibliography
%%
\bibliographystyle{ACM-Reference-Format}
\bibliography{xjurisdiction-divergence}

%%
%% Appendix
%%
\appendix

\section{LLM Judge Prompt}
\label{app:prompt}

\textbf{System:}
\begin{verbatim}
You are a regulatory-affairs expert comparing FDA and EMA
guidance documents. Given an FDA requirement and an EMA
requirement on the same topic, classify their relationship.

Respond with ONLY a JSON object in this exact format:
{"label": "<AGREE|DIVERGE|SILENT>", "reason": "<one sentence>"}

Definitions:
- AGREE: both jurisdictions require substantively the same thing
- DIVERGE: the requirements directly conflict (one mandates
  what the other forbids or softens)
- SILENT: one jurisdiction does not address what the other
  regulates
\end{verbatim}

\textbf{User:}
\begin{verbatim}
FDA requirement:
{fda_text}

EMA requirement:
{ema_text}

Classify the relationship.
\end{verbatim}

Model: \texttt{claude-haiku-4-5-20251001}. Temperature: default (1.0).
Max tokens: 100. Results cached; each pair scored exactly once.

\section{Graph-RAG Prompts}
\label{app:graph_prompts}

Full prompt text is available in the released code repository
(\texttt{graph\_rag.py}). The pipeline uses two sequential LLM calls per pair.

\textbf{Stage~1 — Triplet Extraction.} Each requirement is converted to a
JSON policy node:
\texttt{\{subject, obligation\_level, requirement, conditions\}},
where \texttt{obligation\_level} $\in$ \{MANDATORY, PROHIBITED, RECOMMENDED,
PERMITTED, SILENT\}.

\textbf{Stage~2 — Graph Classification.} The two policy nodes are compared;
the LLM returns \texttt{\{label, reason\}} with AGREE/DIVERGE/SILENT.
Key calibration: modal-register differences alone (RECOMMENDED vs.\ MANDATORY)
do \textit{not} constitute DIVERGE if the substantive requirement is the same.

\textbf{Hard rules (no LLM call):} MANDATORY $\leftrightarrow$ PROHIBITED
$\Rightarrow$ DIVERGE; both SILENT $\Rightarrow$ SILENT.
Model: \texttt{claude-haiku-4-5-20251001}; Stage~2 max tokens: 300.

\end{document}
